---
title: "P8105 Homework 5"
author: "Alison Elgass"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(cache = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
set.seed(10)
```

# Problem 1
```{r}
iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))

fill_missing = function(vect) {
  for (i in 1:length(vect)) { #loop thru vect
    if (is.na(vect[i])) {     #if missing value found
     
      if (is.numeric(vect)) {         #case 1 - numeric
        non_miss = vect[!is.na(vect)] #select non-missing values
        vect[i] = mean(non_miss)      #set missing to mean
      } 
      
      else if (is.character(vect)) { #case 2 - char
        vect[i] = "virginica"        #set missing
      }
    }
  }
  vect
}

iris_filled = vector("list", length = 5) #empty list of 5
iris_filled = map(.x = iris_with_missing, ~fill_missing(.x))
```

# Problem 2
```{r }
output = vector("list", 20)

#list out 20 full data file names
file_names = list.files(path = "./data", full.names = TRUE)

#iterate over file names, read in each
output = map(.x = file_names, ~read_csv(file = .x))

#now convert into a dataframe
trial_data = bind_rows(output) %>% 
  #add columns for subject #, study arm
  mutate(
    subject = row_number(),
    arm = if_else(
      row_number() <= 10, "control", "experimental")
  ) %>% 
  #re-arrange so subject and arm come first
  select(subject, arm, everything())

head(trial_data) #wide format

#now pivot to long format and make a spaghetti plot
trial_data %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    names_prefix = "week_",
    values_to = "value") %>% 
  ggplot(aes(x = week, y = value)) +
  geom_line(aes(group = subject, color = arm))
    
```

The experimental group subjects seem to increase their values over the 8 week study, while the control group generally stays within the same range and does not show a linear trend.  


# Problem 3
First we create a function that takes set parameters and runs a linear regression and outputs the estimate for $\beta$~1~ and the p-value for a test of H~0~: $\beta$~1~ = 0  
  
Then we run the simulation 10,000 times each for $\beta$~1~ = 1-6, and save the values in a long tibble `sim_results`
```{r}
sim_regression = function(n = 30, beta0 = 2, beta1 = 0, sigma2 = 50) {
  
  sim_data = tibble(
    x = rnorm(n, mean = 0, sd = 1),
    y = beta0 + beta1*x + rnorm(n, 0, sigma2) 
  )
  
  ls_fit = lm(y ~ x, data = sim_data)
  
  broom::tidy(ls_fit, conf.level = 0.95) %>% 
    select(estimate, p.value) %>% 
    slice(2) #only second row (beta1 slope)
}

sim_results = 
  tibble(
    beta_list = 1:6
    ) %>% 
  mutate(
    #for each entry in column sample size,
    #rerun SLR 10,000 times, using specified beta
    output_list = 
      map(.x = beta_list, ~ rerun(10000, sim_regression(beta1 = .x))),
    output_df = 
      map(output_list, bind_rows), #list to df
  ) %>% 
  select(-output_list) %>% 
  unnest(output_df)
```

### Plot of Power vs. Beta
```{r}
#only data entries where we reject the null
sim_rejects = 
  sim_results %>% 
  filter(p.value < 0.05)

sim_rejects %>% 
  group_by(beta_list) %>%
  summarize(
    power = n()/10000          #power = # rejections/total
  ) %>% 
  ggplot(aes(x = beta_list, y = power)) + geom_point() +
  xlab("Beta 1 True Value") + ylab("Power")
  
```
Power increases as effect size increases. In this example, effect size is the true value of $\beta$~1~ - the value of $\beta$~1~ in the null hypothesis, which here is 0. This makes sense because as the true value of $\beta$~1~ increases it gets further from the null value so we should correctly reject the null more often.

### Plot of Estimated vs. Actual Beta
```{r}
sim_results %>% 
  group_by(beta_list) %>% 
  summarize(
    avg_beta = mean(estimate)
  ) %>% 
  ggplot(aes(x = beta_list, y = avg_beta)) + geom_point() +
  xlab("True Beta 1") + ylab("Average Beta 1 Estimate")

sim_rejects %>% 
  group_by(beta_list) %>% 
  summarize(
    avg_beta = mean(estimate)
  ) %>% 
  ggplot(aes(x = beta_list, y = avg_beta)) + geom_point() +
  xlab("True Beta 1") + ylab("Average Beta 1 Estimate")
```

The second plot shows the average value of the estimated $\beta$~1~ vs. the actual $\beta$~1~ for the `r nrow(sim_rejects)` simulations where we would reject the null. Unlike the first plot, which considers all the simulations, these estimates are not very close to the actual $\beta$~1~ values. This is because rejecting the null means that the estimate for $\beta$~1~ is not close to the null value of 0, so these are outlier values for $\beta$~1~ which may be very high.
